---
title: "Gasbarro Capstone Project: Opportunities Analysis - Statistics - Section 5"
output:
  md_document: default
  ioslides_presentation: default
  pdf_document: default
  slidy_presentation: default
  word_document: default
  html_document: default
date: "November 14, 2018"
params:
  symbol: Not Applicable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
## IMPORTING THE SPREADSHEET

library(readxl)
library(tidyverse)
library(dplyr)
library(readxl)
library(stringr)
library(lubridate)
library(ggplot2)
library(xts)
library(scales)

opps_revised <- readRDS("opps_revised.rds")

# The following filters are actually putting "tibble" summaries on the output.
# How do we prevent that from happening.
# opps_revised %>% filter(expected_revenue_usd > 100000000)
# opps_revised %>% filter(forecasted_amount_usd > 100000000)


```
# STAGE: Count of Forecasted Dollar Amount (USD)
This first plot shows the count (number of deals) and where they fall in a range of US dollars ($) for the projected Forecasted Dollar Amount (what the deal was proposed to the customer for). As you can see, the vast majority of them fall in a range between USD 500 and USD 1 million, with the concentration being between USD 1,000 and 900,000.

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(opps_revised, aes(x = forecasted_amount_usd)) + 
  geom_histogram() +
  scale_x_log10(labels=dollar)
```

# STAGE: Count of Expected Revenue Amount (USD)
The Expected Revenue is the amount of money we expect to make from the the deal when it closes and commits. A deal might be forecasted for USD 500,000, and the expected revenue might be USD 250,000 (meaning there are USD 250,000 of costs).

The bar plot below for Expected should align with the plot above for Forecasted. Are there noticeable differences in the shape or outline of these plots? If they are closely match, that indicates that expected revenue is aligned with the forecasted amount (meaning, costs are controlled proportionate to the deal). If there is significant difference, this would suggst that there is a misalignment between what was Forecasted and what we Expect to make. For example: If Expected Revenue looks significantly lower, that might indicate that the costs to staff the project increased, meaning our internal costs consumed more money, and the expected revenue decreased proprotionately.


```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
ggplot(opps_revised, aes(x = expected_revenue_usd)) + 
  geom_histogram() +
  scale_x_log10(labels=dollar)

```

## Is there a way to mount one graph on the other?

# STAGE: Count by Year 
One of the original questions our group asked is, "What is overall number of deals we WON AND CLOSED vs. DROPPED and LOST?" 

To answer this question, plot a bar graph that shows the 5 categories in Stage (simple_stage) and the number of deals in each category. As you can see in the chart below, only 3 of the categories (WON, LOST, WIP) have the majority of the deals. 
```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(opps_revised, aes(x = stage_simple)) + 
  geom_bar()

```

Then we wanted to know, "What is overall number of deals WON AND CLOSED, vs. DROPPED and LOST, grouped by YEAR?" 

Below are two bar graphs showing the same information. 

1. The first chart simply adds a fill to the previous chart (in which the Stage is our baseline X axis). This clearly shows 2017 had the highest counts of WON and LOST deals, more than 2016 or 2018.

2. The first chart changes the X asis to the year, and instead uses the fill to identify the stage. This makes it less easier to quickly identify the year which had the highest counts, but does read nicely, and might be used for other charts (e.g. Industry).

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(opps_revised, aes(x = stage_simple, fill=fiscal_year)) + 
  geom_bar()

ggplot(opps_revised, aes(x = fiscal_year, fill=stage_simple)) + 
  geom_bar()


```

# STAGE: Count by Industry

We use the last bar graph as the basis for looking at the count by Stage across the industries. However, this chart is nearly useless: "Everything Else" dominates, while our areas of focus (Distribution, Engineering, Utilities) reduced to very small amounts.


```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(opps_revised, aes(x = industry_simple)) + 
  geom_bar()

ggplot(opps_revised, aes(x = industry_simple, fill=stage_simple)) + 
  geom_bar()


```



# Update Industry Simple with Energy

It also appears we missed something in our additional cleaning operation: Where is the "Energy" sector? Turns out, we lumped it with "Everything Else" incorrectly. So we'll perform an updated operation and refresh our graphs.

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
opps_revised <- opps_revised %>%
  mutate(
    industry_simple = case_when(
      industry == "Aerospace & Defense" ~ "Everything Else",
      industry == "Automotive" ~ "Everything Else",
      industry == "Computer Hardware" ~ "Everything Else",
      industry == "Computer Software" ~ "Everything Else",
      industry == "Consumer Packaged Goods" ~ "Everything Else",
      industry == "Distribution & Transportation" ~ "Distribution",
      industry == "Education" ~ "Everything Else",
      industry == "Energy" ~ "Energy",
      industry == "Engineering & Construction" ~ "Engineering",
      industry == "Federal Public Sector" ~ "Everything Else",
      industry == "Financial Services" ~ "Everything Else",
      industry == "Healthcare" ~ "Everything Else",
      industry == "Hospitals / Health Care" ~ "Everything Else",
      industry == "Insurance" ~ "Everything Else",
      industry == "Legal" ~ "Everything Else",
      industry == "Manufacturing" ~ "Everything Else",
      industry == "Media & Entertainment" ~ "Everything Else",
      industry == "Other" ~ "Everything Else",
      industry == "Pharmaceutical" ~ "Everything Else",
      industry == "Professional Services" ~ "Everything Else",
      industry == "Public Administration" ~ "Everything Else",
      industry == "Public Sector" ~ "Everything Else",
      industry == "Retail" ~ "Everything Else",
      industry == "Telecommunications" ~ "Everything Else",
      industry == "Transportation & Public Utilities" ~ "Transportation",
      industry == "Travel" ~ "Everything Else",
      industry == "Utilities" ~ "Utilities",
      industry == "" ~ "Missing"
    )
  )

saveRDS(opps_revised,file="opps_revised.rds")

ggplot(opps_revised, aes(x = industry_simple)) + 
  geom_bar()

ggplot(opps_revised, aes(x = industry_simple, fill=stage_simple)) + 
  geom_bar()

```

Now let's filter out "Everything Else" and concentrate on the Industries we selected: "Energy" , "Utilities", "Engineering","Distribution"

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(subset(opps_revised,industry_simple %in% c("Energy" , "Utilities", "Engineering","Distribution")), aes(x = industry_simple)) + 
  geom_bar()

ggplot(subset(opps_revised,industry_simple %in% c("Energy" , "Utilities", "Engineering","Distribution")), aes(x = industry_simple, fill=stage_simple)) + 
  geom_bar()

```

Energy (which comprises Oil, Gas, Fossil and Mineral industries) and Utilities are the sectors with the most deals.

What if we break down by year?

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(subset(opps_revised,industry_simple %in% c("Energy" , "Utilities", "Engineering","Distribution")), aes(x = industry_simple, fill=fiscal_year)) + 
  geom_bar()


```

# Section 5.3 Questions

The following are a list of questions from Section 5.3 exercise, which will be posted to Github.

1. Can you count something interesting?
As seen the charts abobe, we counted a number of interesting facts on the amount of deals by Forecasted Amount and Expected Revenue, by Stage, and by Industry.

2. Can you find some trends (high, low, increase, decrease, anomalies)?
We noticed the following
* Forecasted Amount and Expected Revenue seemed to be in alignment. (Can we statistically say they are correlated? I'm unsure about this.) 
* The Year 2017 saw the highest count of WON and LOST deals than any other year.
* The Energy and Utilities sector also saw higher deal activities than any other sectors we're tracking (Distribution, Engineering). 

3. Can you make a bar plot or a histogram?
As demonstrated above, we made a lot of bar plots and histograms. We might try doing other graphics in the subsequent demonstrations.

4. Can you compare two related quantities?
Yes and No. The bar graphs above do have fills that allow us to compare data for Stages, Industries and Year (depending on the configuration). However, I was unable to figure out a way to compare the Forecasted Amount and Expected Revenue by putting one graph on top of the other.

5. Can you make a scatterplot?
I can make a scatterplot. However, I'm not sure that the volume of data in the Opportunities datafram (over 200K) lends itself to this kind of visual display, not without significant filtering. The bar charts appear to be better at group large volumes of records and variables in a visually pleasing way.

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

# Basic scatter plot
ggplot(opps_revised, aes(x=industry_simple, y=expected_revenue)) + 
  geom_point()

ggplot(opps_revised, aes(x=industry_simple, y=expected_revenue)) + 
  geom_point(size=2, shape=23)

```

6. Can you make a time-series plot?
The Date/Time values in the Opportunities dataframe do not seem to lend themselves to a time-series plot. Each record only captures a single Date/Time variable for the Opportunity - the Create Date and the Close Date. It is not tracking any series of activities for the Opportunity except these 2 dates.


# What's Next?
Section 5.3 Asks: Having made these plots, what are some insights you get from them? Do you see any correlations? Is there a hypothesis you would like to investigate further? What other questions do they lead you to ask?

Some of my ideas for Section 6 exercises and beyond.

1. Graphics beyond bar charts, or advanced bar charts (instead of one bar with multiple fills, multiple bars grouped together with different fills)

2. Comparisons and correlations for the following variables
* Lead Source
* Currency
* Territory - this could be a new variable, extrapolated from Currency (by as Nation instead of Dollar Symbol) - will require additional change to dataframe

3. I think there are more combinatorials to be tried - Dollar Amounts vs. Stage, Dollar Amounts vs. Industry, Lead Source vs. Industry

4. Trying to visualize 3 or 4 sets of variables in the same graph. How is this possible?
* Example: For each Industry, a break down by Year of the number of deals WON and LOST, and the Total Dollar Amounts (Forecasted and/or Expect) for WON and LOST.


