---
title: 'Gasbarro - Linear Regression - Exercise 0'
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
date: "February 18, 2019"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 0: Linear Regression

## Load the STATES data

```{r, 0.1, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
## IMPORTING THE DATA

library(readxl)
library(tidyverse)
library(dplyr)
library(readxl)
library(stringr)
library(lubridate)
library(ggplot2)
library(xts)
library(scales)

# Load the data
statesData <- readRDS("dataSets/states.rds")

```

## Examine and plot the data before fitting the model
```{r, 0.2, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

# Select data for METRO areas and ENERGY consumption
statesMetroEnergy <- subset(statesData, select = c("metro", "energy"))

summary(statesMetroEnergy)

# Plot X=METRO against Y=ENERGY
ggplot(statesMetroEnergy, aes(x=metro,y=energy)) +
  geom_point()
```

## Print and interpret the model
```{r, 0.3, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

cor(statesMetroEnergy, use="pairwise")

modelMetroEnergy <- lm(energy ~ metro, data = statesMetroEnergy)

summary(modelMetroEnergy)

```

## Plot the model to look for deviations from modeling assumptions
```{r, DATAPREP, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

#Assuming we use the lm dataframe we just generated

plot(modelMetroEnergy)

# CAN YOU USE GGPLOT TO GENERATE THE SAME PLOTS?

```

## Select one or more additional predictors 
Select one or more additional predictors to add to your model and repeat steps 1-3. Is this model significantly better than the model with metro as the only predictor?

For this exercise, we're using DENSITY and INCOME

```{r, 0.4, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

# Select data for METRO, ENERGY, DENSITY, and INCOME
statesMetEnerDenInc <- subset(statesData, select = c("metro", "energy", "density", "income"))

summary(statesMetEnerDenInc)

# Plot X=METRO against Y=ENERGY
ggplot(statesMetEnerDenInc, aes(x=density,y=energy)) +
  geom_point()

ggplot(statesMetEnerDenInc, aes(x=income,y=energy)) +
  geom_point()

cor(statesMetEnerDenInc, use="pairwise")

modelDenEner <- lm(energy ~ density, data = statesMetEnerDenInc)

summary(modelDenEner)

plot(modelDenEner)
```

# Exercise 1

## 1.1 Add on to the regression 
Add on to the regression equation that you created in exercise 1 by generating an interaction term and testing the interaction.

I used the following variables in my previous exercise:
* Metro
* Energy
* Density
* Income

```{r, 1.1, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

# Use statesMetEnerDenInc
#modelDenEner <- lm(energy ~ density, data = statesMetEnerDenInc)

modelEnergyMetroXDensity <- lm(energy ~ metro * density, data = statesMetEnerDenInc)

summary(modelEnergyMetroXDensity)

plot(modelEnergyMetroXDensity)

```


## 1.2 Regions
Try adding a region to the model. Are there significant differences across the four regions?

? Any easier way to do naming conventions?

```{r, 1.2, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

# Select data for METRO, ENERGY, DENSITY, INCOME, REGIONS
statesMetEnerDenIncReg <- subset(statesData, select = c("metro", "energy", "density", "income","region"))

modelEnergyMetroXDensityRegion <- lm(energy ~ metro * density + region, data = statesMetEnerDenIncReg)

# I think we're supposed to use ANOVA for this exercise
anova(modelEnergyMetroXDensityRegion)


```

# Exercise 2

Load the NatHealth2011.rds data

```{r, 2.0, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

NatHealth2011 <- readRDS("dataSets/NatHealth2011.rds")
labs <- attributes(NatHealth2011)$labels

```

## 2.1 Use GLM
Use glm to conduct a logistic regression to predict ever worked (everwrk) using age (age_p) and marital status (r_maritl).

```{r, 2.1, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

# Select the 3 variables
NH2011_WorkAgeMar <- subset(NatHealth2011, select = c("everwrk", "r_maritl", "age_p"))

# Summary before transformation
summary(NH2011_WorkAgeMar)

# Transform using a factor on Work, and drop any unused levels in the Marital Status
NH2011_WorkAgeMar <- transform(NH2011_WorkAgeMar, everwrk = factor(everwrk, levels = c("1 Yes", "2 No")), r_maritl = droplevels(r_maritl))

Model_NH2011_WorkAgeMar <- glm(everwrk ~ age_p + r_maritl, data = NH2011_WorkAgeMar, family = "binomial")

summary(Model_NH2011_WorkAgeMar)

```

#2.2 Working v Marital Prediction
Predict the probability of working for each level of marital status.

```{r, 2.2, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

# Install and Load the Effects library

#install.packages("effects")
library(effects)

# Use Effect to predict the probability on the model from the previous exercise

data.frame(Effect("r_maritl", Model_NH2011_WorkAgeMar))
 
```


# Exercise 3

Where's the bh1996 dataset?

```{r, 3.0, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

# Need to load a package called MULTILEVEL
#install.packages("multilevel")
data(bh1996, package="multilevel")
 
```
